Parsing Postfix log files

### 2009-03-17 doc/logparser-implementation.tex
Wrote the framework section of the implementation chapter.

XXX TO BE WRITTEN\@.

XXX GENERIC RULES\@.

XXX COMPRESSED FILES\@.

XXX STATE TABLES\@.

XXX LOAD AND VERIFY RULES\@.

@@

XXX HALF OF THIS CONTENT IS IN FOOTNOTES, WHICH SHOULD BE INLINED\@.

@@

@@

@@

\newpage{} % XXX ADDED BECAUSE OF FLIP-FLOPPING LAYOUT.

>>>>

The role of the framework in this architecture was described in detail in
\sectionref{framework in architecture}; this section is concerned with the
implementation of the framework within \parsername{}.  The framework
manages the parsing process, taking care of the drudgery and boring tasks,
provides services to the actions, and implements some efficiency measures.

Each time the parser is run the framework performs some initialisation
tasks, setting up data structures that will later be used by actions,
including several state tables.  Data about the connections and mails being
processed is held in \texttt{connections} and \texttt{queueids}
respectively.  There are also state tables used when working around
complications in parsing: \texttt{timeout\_queueids} is used when dealing
with connections that time out during the DATA phase (\sectionref{timeouts
during data phase}); \texttt{bounce\_queueids} is part of the solution to
bounce notification mails being delivered before their creation is logged
(\sectionref{Bounce notification mails delivered before their creation is
logged}); and \texttt{postsuper\_deleted\_queueids} caches information
about mails recently deleted by the administrator so that subsequent log
lines processed by the SAVE\_DATA action can be discarded (\sectionref{Mail
deleted before delivery is attempted}).

When the framework loads the ruleset, it verifies each of the rules,
checking:

    \item That the specified action is registered with the framework.

    \item Ensuring the regex is valid and compiling it for efficiency
        (\sectionref{Caching compiled regexes}).

    \item For overlap between the data captured by the regex and additional
        data specified in either \texttt{connection\_data} or
        \texttt{result\_data}.

    \item That \texttt{connection\_data}, \texttt{result\_data}, and the
        regex captures specify valid fields to save data to.

If there are state tables from a previous \parsername{} run to be loaded,
they will be loaded now; obviously the framework supports saving its state
tables too.  The need to track re-injected mail (\sectionref{Re-injected
mails}) complicates the process of saving state, as the relationships
between mails must be maintained; loading state is complicated by dealing
with aborted delivery attempts (\sectionref{aborted delivery attempts}),
because a different set of relationships between connections and mails must
be re-created when the state tables are restored.

The framework will now move on to parsing each log file.  For each log line
it will use those rules whose \texttt{program} field matches the program in
the log line (see \sectionref{rule conditions in implementation}), falling
back to generic rules if necessary, and warning if the log line is
unrecognised.  The rules are sorted for increased efficiency; see
\sectionref{rule ordering for efficiency} for details.  The repetitive
nature of log files gives them high compression ratios; the framework
supports reading compressed files, simplifying the process of parsing them
because they do not have to be uncompressed to a temporary file before
parsing begins.  The framework also displays a progress bar to show how far
parsing has progressed through the log file, and to give the user an
indication of how long the remainder of the parsing process is likely to
take.  The progress bar is not as reliable when parsing compressed files
because a compressed block may uncompress to a variable number of log
lines; there is also variation between the recognition and processing time
of individual log lines, but overall the progress bar is a useful addition.

The framework collects data used when evaluating \parsernames{} efficiency
for the results chapter (\sectionref{Results}); in addition some of the
techniques used to improve parsing speed can be turned off or altered to
show the effect they have.  Three sets of data are collected:

    \item How long it takes to parse each log file.

    \item The number of rules tried when recognising log lines; the
        framework may need to try multiple rules for a given log line
        before it finds one that recognises it.

    \item How many log lines did the framework attempt to recognise, how
        many were skipped because there were no rules whose condition
        matched that log line's program, how many were successfully
        recognised, and how many were not recognised.

There are five ways the framework can adapt its behaviour to demonstrate
how effective the optimisations it uses are, and how efficient
\parsername{} is:

    \item The rule ordering used can be changed from optimal (the default,
        most efficient) to shuffled (intended to represent an unordered
        ruleset) or reverse (reverse of optimal, least efficient).  The
        results of this are shown in \sectionref{rule ordering for
        efficiency}.

    \item The framework can record which rule recognised each log line, and
        then on a subsequent run use that information so that it uses the
        correct rule for each log line.  This gives the best possible
        running time, as only one rule is used for each log line.  A
        variation of this is to use the correct rule last, to get the worst
        possible running time.  \sectionref{perfect rule ordering}
        discusses this in detail.

    \item Each regex is compiled and the result cached when the rules are
        loaded; this can be disabled and the regex compiled every time it
        is used when recognising log lines, to demonstrate the effect this
        optimisation has on parser efficiency, as shown in
        \sectionref{Caching compiled regexes}.

    \item Normally once a log line has been recognised, the framework
        invokes the action specified by the rule.  Invoking the action can
        be skipped if desired, so the timing information shows how long is
        spent recognising log lines; this time can be subtracted from the
        time taken by a normal run to show how long is spent processing log
        lines.  This data is analysed in \sectionref{recognising vs
        processing}.

    \item The framework can skip inserting data into the database, because
        that dominates the execution time of the parser, and the evaluation
        in \sectionref{Results} is measuring the speed of \parsername{},
        not the speed of the database or the disks it resides on.

The framework also provides several debugging options, to aid in writing or
correcting rules, or figuring out why the parser is not behaving as
expected.  In increasing order of severity they are:

    \item The regex from the recognising rule and the log line it
        recognised can be printed, so that the user can verify that the
        correct rule recognised each log line.  A possible variant of this
        would be to mark specific rules so only their regexes and the log
        lines they recognise would be printed.

    \item Each result can be extended with extra debugging information, so
        that when a mail's data is dumped for inspection more information
        is available.  The information currently added is: the log line's
        timestamp in human readable form; the entire log line; the name of
        the log file and the number of the log line within it.

    \item Every connection and result added to the database can be dumped
        in a human readable form.  This will result in a huge amount of
        debugging information, so it is only useful for small log file
        snippets, because otherwise the amount of information is
        overwhelming.

@@

XXX HALF OF THIS CONTENT IS IN FOOTNOTES, WHICH SHOULD BE INLINED\@; IT
ALSO NEEDS TO BE PROOFREAD\@.

@@

\label{rule conditions in implementation}

@@

XXX MENTION RULE CONDITIONS HERE\@; STRIP THE EXISTING STUFF FROM THE RULES
TABLE AND EXAMPLE RULE, THEN HAVE THEM REFER TO HERE\@.

@@

### 2009-03-17 doc/logparser-results.tex
Added a diagram of the database schema.  It took a long time to find a tool to draw it with, longer to draw it, and even longer still to make it fit on the page properly.
Checked all references in the implementation chapter.

Finished the flow chart introduction.

Some minor changes and corrections.
Wrote the framework section of the implementation chapter.

XXX COMPARE SAVING RESULTS TO RAM DISK VERSUS SAVING TO HARD DISK\@.

@@

@@

>>>>

@@

XXX DO A TEST-SOME RUN AND SAVE THE RESULTS TO THE DATABASE, THEN COMPARE
THE TIME WITH A RUN THAT DOES NOT SAVE RESULTS\@; PUT THE COMPARISON IN THE
INTRODUCTION, NEAR THE MACHINE SPECS\@.

@@

\subsection{Perfect Rule Ordering}

\label{perfect rule ordering}

XXX WRITE ABOUT PERFECT RULE ORDERING\@; DUNNO WHETHER TO PUT IT IN WITH
RULE ORDERING FOR EFFICIENCY, OR IN ITS OWN SECTION\@.

\subsection{Where Is The Time Spent: Recognising Or Processing Log Lines?}

\label{recognising vs processing}

XXX EXAMINE THE TIME SPENT RECOGNISING LOG LINES VS THE TIME SPENT
PROCESSING THEM\@.
